Risks of Running Vulnerable Container Images for Data Analysis on HPC Clusters

### 2019-11-07 paper.tex
added results

%%%% Packages %%%%
\usepackage{siunitx}
\usepackage{algpseudocode} % Algorithmic environment
\usepackage{xspace}
\usepackage{minted} % Used for JSON highlighting
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{makecell}
\usepackage[flushleft]{threeparttable}

@@

\author[1]{Bhupinder Kaur}
\author[1]{Aiman Hanna}
\author[1]{Tristan Glatard}

@@

Containers, which provide operating system (OS) level virtualization, are very
popular. This is because they are very lightweight, flexible, resource-efficient,
provide fast development cycles, continuous delivery,
and cost reduction of infrastructure. OS-level virtualization gained a lot of interest when Docker was
introduced in 2013~\cite{gantikow2016providing}. However, it has security concerns due to which it
is not supported by High Performance
Computing (HPC) environments. As a result, Singularity came in 2016 with the goal of providing support
for HPC environments. It is used in many clusters including Compute Canada.
Table \ref{fig:cluster_list} illustrates the reported usage of Singularity on HPC clusters~\cite{kurtzer2017singularity}.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{Figures/cluster_list.png}
	\caption{List of HPC clusters using Singularity~\cite{kurtzer2017singularity}.}
	\label{fig:cluster_list}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=.7\columnwidth]{Figures/container.png}
  \caption{Architecture of Container-based
                Virtualization.}
  \label{fig:container-overview}
\end{figure}

\textit{Privilege Escalation} A malicious attacker can break out of the container
and gain control of the host system and other containers running on the same host.

@@

However, two features of the Linux kernel, \textit{control groups} and \textit{namespaces},
are used by containers to mitigate these threats to a certain extent. Security
of containers is still most concerning issue. According to a Forrester survey~\cite{} done in January
2015, security is the
biggest concern while deploying containers.
Except web applications, containers are also popularly used for scientific computing
and data science in neuroscience. In fact, Singularity was developed for scientific applications
by keeping focus on HPC environments. This popularity of containers lead us to the
primary research question of this work: \textit{what is the current security state of
container images that are deployed on HPC clusters}.

To answer that question, we scan all images that are used on Compute
Canada clusters by neuroscience field. Scanning refers to the practice of
collecting all information about the container image, investigating it
for vulnerabilities, and finally producing a
report to summarize scanning results. It is broadly categorized into
two types: static scanning and dynamic scanning. The former scanning technique
involves extracting information of an image without running it into a container,
whereas the latter approach involves launching the container from the target image, and
then scanning the running container.
Here, we focus mainly on vulnerability scanning and not on malware.
Malware is a computer software that is specially designed to attack the target computer,
whereas vulnerability is a weakness in the developer's software that can be used
by the attacker to perform unathorized actions on the target computer.
In this paper, we report all vulnerabilities that are present in
container images, specifically used by neuroimaging field, and access their
related risks for data analysis. Particularly, we aim to answer three primary
research questions:

@@

To answer those questions, we take an example of neuroscience field where
containers are used on Compute Canada HPC cluster.
We are not the first to work on vulnerabilities inside container images but
the existing research work on containers focuses mainly on the security of Docker
containers.
This focus is justified by the fact that, containers expose the host's resources
(e.g., file system/ IPC) to the guest system. This feature raises a confidentiality
threat for the applications running on the same host. Previous studies evaluated
security of Docker engine ~\cite{martin2018docker, sultan2019container, combe2016docker, bui2015analysis},
and have scanned vulnerabilities on Docker hub~\cite{Shu2017, gummaraju2015over}.
Also Singularity came up with Stools, which can be used for scanning Singularity images
for security and quality checks.
Here, we focus on the specific context of scientific data analysis on HPC clusters, taking
neuroimaging as an example.

@@

\subsection{Container Engine: Singularity}

@@

BIDS apps~\cite{gorgolewski2017bids} is a framework designed to share and execute neuroimaging
analysis pipelines. It improves usability, accesibility, and reproducibility
of different applications through containerization.
Currently, there are 27 BIDS apps available and all are using different Docker images.

\begin{figure}
  \centering
	\includegraphics[width=\columnwidth]{Figures/bids_table.png}
  \caption{All Container images used by BIDS apps.}
  \label{fig:bids_apps}
\end{figure}

\subsubsection{ReproNim}

@@

@@

If there exist vulnerabilities in the container image, then there are two ways
to proceed, either not to run that image at all, or run only after updating the
image. Updating the image involves updating all software packages that are
inside that image. However, this option effects the reproducibility
of the image, which means that the image cannot be used to verify research
findings of other scientists. An alternative approach can be to trim unnecessary
packages from images, which in turn reduces the number of vulnerabilities
present in images. To update container images we used a bash script
which is given in Figure~\ref{}.

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=0pt,
               tabsize=4]{bash}
#!/usr/bin/env bash

function die {
	echo $1
	exit 1
}

function find-pm {
		(type $1 &> /dev/null && echo "> Found $1") || \
		(echo "> $1 not found" && return 1)
}

\rm -f .updated
find-pm yum && (yum update -y && touch .updated) 
find-pm apt && (apt update -y && apt upgrade -y \
&& apt autoremove -y && apt autoclean -y \
&& touch .updated)
test -f .updated || die "Cannot find package manager \
or update failed"	       
\end{minted}
\caption{update.sh}
\label{listing:update.sh}
\end{listing}

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=0pt,
	       tabsize=4]{bash}
#!/usr/bin/env bash

set -ue

function usage {
	echo "$0 <image_list>"
	exit 1
}

if [ $# -lt 1 ]
then
	usage
fi

for image in $*
do
	echo -n "Updating image ${image}... "
	docker run -e /bin/bash -v $PWD:$PWD -w \
	$PWD ${image} ./update.sh &>${image}.log \
	&& echo "[ OK ]"
	echo -n "Commiting updated image... "
	ID=$(docker ps -a | head -2 | awk \
	'$1!="CONTAINER" {print $1}')
	docker commit ${ID} updated_${image} \
	&& echo "[ OK ]"
done
\end{minted}
\caption{update-images.sh}
\label{listing:update-images.sh}
\end{listing}

@@

Anchore is an end-to-end, open-source container security platform. Anchore
is a static
scanning tool that analyses container images and lists OS
packages, non-OS packages (Python, Java, Gem, and Npm), and files.
Anchore is a set of tools, which users can
employ for scanning container images and applying custom policies to them.
It provides visibility and transparency of the container environment.
Anchore offers services in two ways: Anchore Engine and Anchore Engine
Command Line Interface (CLI). Anchore Engine is available as a Docker
image, which is available on Docker Hub. Anchore Engine CLI provides
command line interface in addition to Anchore Engine REST API. Anchore
Engine CLI can be installed directly on the host and used to inspect
images.

\begin{figure}
  \centering
  \includegraphics[width=.7\columnwidth]{Figures/anchore.png}
  \caption{Databases used by Anchore.}
        \label{fig:anchore_db}
\end{figure}

\TG{You should refer to the vulnerability databases used, and explain how
software components, for instance files, are matched to these databases. Is
the matching based on file checksum?}

@@

\begin{figure}
  \centering
  \includegraphics[width=.7\columnwidth]{Figures/vuls.png}
  \caption{Databases used by Vuls.}
	\label{fig:vuls_db}
\end{figure}

@@

%\bibliographystyle{vancouver-authoryear}
\bibliography{bibliography}

>>>>

%%%% Packages %%%%
\usepackage{siunitx}
\usepackage{algpseudocode} % Algorithmic environment
\usepackage{xspace}
\usepackage{csvsimple}
\usepackage{minted} % Used for JSON highlighting
\usepackage{datatool}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{caption}
\usepackage[justification=centering]{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{adjustbox}
\usepackage{makecell}
\usepackage[flushleft]{threeparttable}

@@

\author[1]{Bhupinder Kaur}
\author[1]{Tristan Glatard}
\author[1]{Aiman Hanna}

@@

Containers, which provide operating system (OS) level virtualization, are very
popular. This is because they are very lightweight, flexible, resource-efficient,
provide fast development cycles, continuous delivery,
and cost reduction of infrastructure. OS-level virtualization gained a lot of interest when Docker was
introduced in 2013~\cite{gantikow2016providing}. However, Docker has security concerns due to which it
is not supported by High Performance
Computing (HPC) environments. As a result, Singularity came in 2016 with the goal of providing support
for HPC environments. It is used in many clusters including Compute Canada.
Table \ref{fig:cluster_list} illustrates the reported usage of Singularity on HPC clusters~\cite{kurtzer2017singularity}.

\begin{comment}
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{Figures/cluster_list.png}
	\caption{List of HPC clusters using Singularity~\cite{kurtzer2017singularity}.}
	\label{fig:cluster_list}
\end{figure}
\end{comment}

@@

\begin{comment}
\begin{figure}
  \centering
  \includegraphics[width=.7\columnwidth]{Figures/container.png}
  \caption{Architecture of Container-based
                Virtualization.}
  \label{fig:container-overview}
\end{figure}
\end{comment}
\textit{Privilege Escalation} A malicious attacker can break out of the container
and gain control of the host system and other containers running on the same host.

@@

However, two features of the Linux kernel, \textit{control groups} and \textit{namespaces},
are used by containers to mitigate these threats to a certain extent. Security
of containers is still most concerning issue. According to a Forrester survey~\cite{bettini2015vulnerability}
done in January 2015, security is the
biggest concern while deploying containers.
Except web applications, containers are also popularly used for scientific computing
and data science in neuroscience. In fact, Singularity was developed for scientific applications
by keeping focus on HPC environments. This popularity of containers lead us to the
primary research question of this work: \textit{what is the current security state of
container images that are deployed on HPC clusters}.

To answer that question, we scan all images that are used on Compute
Canada clusters by neuroscience field. Scanning refers to the practice of
collecting all information about the container image, investigating it
for vulnerabilities, and finally producing a
report to summarize scanning results. It is broadly categorized into
two types: static scanning and dynamic scanning. The former scanning technique
involves extracting information of an image without running it into a container,
whereas the latter approach involves launching the container from the target image, and
then scanning the running container.
Here, we focus mainly on vulnerability scanning.
\begin{comment}
Malware is a computer software that is specially designed to attack the target computer,
whereas 
\end{comment}
Vulnerability is a weakness in the developer's software that can be used
by the attacker to perform unathorized actions on the target computer.
In this paper, we report all vulnerabilities that are present in
container images, specifically used by neuroimaging field, and access their
related risks for data analysis. Particularly, we aim to answer three primary
research questions:

@@

To answer those questions, we take an example of neuroscience field where
containers are used on Compute Canada HPC cluster.
The existing research work on containers focuses mainly on the security of Docker
containers.
This focus is justified by the fact that, containers expose the host's resources
(e.g., file system/ IPC) to the guest system. This feature raises a confidentiality
threat for the applications running on the same host. Previous studies evaluated
security of Docker engine ~\cite{martin2018docker, sultan2019container, combe2016docker, bui2015analysis},
and have scanned vulnerabilities on Docker hub images~\cite{Shu2017, gummaraju2015over}.
Also Singularity came up with Stools, which can be used for scanning Singularity images
for security and quality checks.
Here, we focus on the specific context of scientific data analysis on HPC clusters, taking
neuroimaging as an example.

@@

\subsection{Container Engines}

\subsection{Singularity}

@@

BIDS apps~\cite{gorgolewski2017bids} is a framework designed to share and execute neuroimaging
analysis pipelines. It improves usability, accesibility, and reproducibility
of different applications through containerization.
Currently, there are 27 BIDS apps available and all are using different Docker images.
%\begin{figure}
%\csvreader[%
% respect all,%
%  before reading=\footnotesize\sisetup{round-mode=places,round-precision=2,round-integer-to-decimal}
%      %\caption{Baseline GF Competition Circle Track 2014}\label{tab:baselline1}
%      \begin{adjustbox}{max width=\columnwidth},
%  after reading=\end{adjustbox},
% tabular={c |c | c | c | c},
%	table head =Image & {OS Distro} & {Vulnerabilities by Anchore} & {Vulnerabilities by Clair} & {Vulnerabilities by Vuls}\\\hline,
% late after line= \\,
%	late after last line=\\\hline %\multicolumn{5}{c}{Total Score: 1481}
%]{results.csv}{}{\csvlinetotablerow}%
%       \centering
%	\caption{Existing number of Vulnerabilities}
%\end{figure}

\begin{comment}
\subsubsection{ReproNim}

@@

\end{comment}

\subsection{Linux Distributions}

BIDS and Boutiques images that we took for our experiment have diverse Linux
distributions such as Ubuntu, Centos, Debian, and Alpine.
Some Ubuntu distributions that are present inside images are old enough
that Ubuntu has stopped providing support such as 14.04 Long Term Support (LTS), 17.04, and 17.10. These
distributions have reached End Of Life (EOL). They are still used in images, and
as there are no options of getting updates due to EOL, it results in vulnerable images.
However, Extended Security Maintenance 
(ESM) can be requested for Ubuntu 14.04 but it is not feasible to use inside images.

@@

If there exist vulnerabilities in the container image, then there are two ways
to proceed, either not to run that image at all, or run only after updating the
image. Updating the image involves updating all software packages that are
inside that image. However, this option effects the reproducibility
of the image, which means that the image cannot be used to verify research
findings of other scientists. An alternative approach can be to trim unnecessary
packages from images, which in turn reduces the number of vulnerabilities
present in images.

@@

Anchore is an end-to-end, open-source container security platform. Anchore
is a static
scanning tool that analyses container images and lists OS
packages, non-OS packages (Python, Java, Gem, and Npm), and files.
Anchore is a set of tools, which users can
employ for scanning container images and applying custom policies to them.
It provides visibility and transparency of the container environment.
Anchore offers services in two ways: Anchore Engine and Anchore Engine
Command Line Interface (CLI). Anchore Engine is available as a Docker
image, which is available on Docker Hub. Anchore Engine CLI provides
command line interface in addition to Anchore Engine REST API. It can be
installed directly on the host and used to inspect images.
Anchore compares package versions, which are present inside the
image, with the vulnerability database to flag vulnerabilities.

@@

\subsection{Clair Scanner}

CoreOS originally created Clair, which is a static vulnerability scanner for containers which
currently support App Container (appc) and Docker~\cite{coreos_2019}. The main goal of the
tool is to provide a transparent look of container’s security. Consequently, it is
named as Clair, simply implying clear. Clair scans container images in regular
intervals, and stores metadata of vulnerabilities, after getting from configured
set of sources. Clair API can be used by clients to query its database to get
the list of vulnerabilities that a particular image has. Clair sends a notification
to alert systems that there is an updation in the vulnerability metadata. Quay
registery uses Clair internally as the security engine. However, Clair does not 
provide facility of comparing
the vulnerability list of a particular image against a whitelist~\cite{arminc_2019}. So to remove
that drawback, another tool from CoreOS, Clair Scanner can be used. It scans image, prepares the list of
vulnerabilities, compares that list against the whitelist, and flags vulnerabilities
that are not present in whitelist.

@@

\section{Results}

\subsection{Existing Vulnerabilities}

Tables show results of scanning BIDS and Boutiques images with above discussed three scanners.
The number of vulnerabilities reported by these scanners are drastically different from each other.
The question arises why these scanners report different number and type of vulnerabilities.
How their logic differs and what exactly is making these results different than each other.
According to our investigation of these results pricisely, below are some of the reasons of
getting different results from these scanners. These reasons include bugs in scanner 's logic,
some scanners ignoring vulnerabilities in a particular package, referring to different databases,
how frequently they are updating their database, etc,. We have explained these differences below in detail
and have provided an example with each reason to make it clear.

\begin{figure}
\csvreader[%
 respect all,%
  before reading=\footnotesize\sisetup{round-mode=places,round-precision=2,round-integer-to-decimal}
      %\caption{Baseline GF Competition Circle Track 2014}\label{tab:baselline1}
      \begin{adjustbox}{max width=\columnwidth},
  after reading=\end{adjustbox},
 tabular={c |c | c | c | c},
        table head =Image & {OS Distro} & {Vulnerabilities by Anchore} & {Vulnerabilities by Clair} & {Vulnerabilities by Vuls}\\\hline,
 late after line= \\,
        late after last line=\\\hline %\multicolumn{5}{c}{Total Score: 1481}
]{results.csv}{}{\csvlinetotablerow}%
       \centering
        \caption{Existing number of Vulnerabilities}
\end{figure}

\begin{figure}[!htb]
        {\includegraphics[width=\columnwidth]
        {Figures/vulngraph.png}}
        \caption{\label{fig:graph1} Scatterplot representing \#Packages and \#Vulnerabilities}
      \end{figure}

\subsection{Comparisons between Scanners}

\subsection{V-C}

V-C represents extra vulnerabilities that are shown by Vuls than Clair.
So the next question is why they are not shown by Clair then. 
Below are some reasons to answer that question:
\begin{enumerate}

\item\textbf{With Status as “Not-affected”-} It means these vulnerabilities are affecting Ubuntu but they are not 
affecting particular Ubuntu release that is present inside the image. The reason why its not affecting 
that release is also mentioned. For example, one reason can be that code which is creating issue is not 
present in that release etc. For example CVE-2015-2305 is not affecting package cups in Trusty.

\item\textbf{Ignored(reached end-of-life)-} These vulnerabilities are ignored in Ubuntu 14.04 LTS as it reached end-of-life 
but does not exist (status as “DNE”) in Ubuntu 14.04 ESM release. As Clair is only referring Ubuntu 14.04 ESM 
release it do not detect these vulnerabilities. For example CVE-2017-9994.

\item\textbf{Not present in Ubuntu 14.04 ESM release-} These vulnerabilities are not present in Ubuntu 14.04 ESM so they 
are not detected by Clair but they affect Ubuntu 14.04 LTS That’s why detected by Vuls. For example, CVE-2017-1375.

\item\textbf{False negatives by Clair-} Clair is missing vulnerabilities that it should detect. For example CVE-2016-9082 
is present in package cairo in Xenial release and it is not detected by Clair.
\end{enumerate}

\subsection{C-V}

C-V represents extra vulnerabilities that are shown by Clair than Vuls. Following are the reasons to explain that:

\begin{enumerate}
\item \textbf{False positives by Clair-}These vulnerabilities are not linked to Ubuntu 14.04 LTS which is 
present inside the image but Clair is showing because these vulnerabilities are present in Ubuntu 14.04 ESM and 
Clair is only referring to Ubuntu 14.04 ESM release. For example CVE-2019-1274.
\item\textbf{False Negatives by Vuls-} For example CVE-2019-5094.
\end{enumerate}

\subsection{A-C}

A-C represents extra vulnerabilities shown by Anchore than Clair. Below are the reasons:
\begin{enumerate}
	\item Linux Package: Clair is not detecting all vulnerabilities in the source “Linux” package. 
		For example CVE-2019-9506.
	\item Ignored(reached end-of-life): These vulnerabilities are ignored in Ubuntu 14.04 LTS as it 
		reached end-of-life but does not exist (DNE) in ESM release. As Clair is only referring 
		Ubuntu 14.04 ESM release it do not detect it. For example CVE-2017-994
	\item Database difference: Clair is using a Docker image for database. So it is not updated everyday. 
		When I scanned images I used the latest available container. For example CVE-2019-5094 was not 
		updated in Clair container that time.
	\item False negatives by Clair: Clair is missing vulnerabilities that it should detect. For example 
		CVE-2016-9082 is present in package cairo in Xenial release and it is not detected by Clair
\end{enumerate}

\subsection{C-A}
C-A represents the vulnerabilities that are not detected by Anchore. Below are the reasons:
\begin{enumerate}

	\item\textbf{EPOCH bug-} There is a bug in the Anchore logic due to which some vulnerabilities are 
		not detected by Anchore. This bug gets triggered when package’s version have epoch 
		(1:3.3.9-1ubuntu2.3)  in it. For example CVE-2018-1125 in procps package
	\item\textbf{Ignored (out of standard support) bug-} There is another bug in Anchore due to 
		which it is not able to detect some vulnerabilities. Due to this bug Anchore does not detect 
		vulnerabilities which have a status as “ignored (out of standard support)” in Ubuntu 14.04 LTS. 
		But if this vulnerability is present in Ubuntu 14.04 ESM release it is detected by Clair. 
		For example CVE-2019-13565 in Ubuntu 14.04
	\item\textbf{Ignored (reached end-of-life)-}Some vulnerabilities have status as “ignored (reached end-of-life)” 
		so Anchore checks the last status of these vulnerabilities whether OS package was vulnerable to the cve 
		prior to the end-of-life. If it was vulnerable only then Anchore will report it, otherwise not. But this 
		vulnerability is present in Ubuntu 14.04 ESM release Clair is reporting that. For example CVE-2019-9924
\end{enumerate}

\subsection{A-V}

A-V represents the vulnerabilities that are detected by Anchore but not detected by Vuls. Below are the reasons 
of these differences:
\begin{enumerate}

	\item\textbf{Linux Package-} Vuls is not detecting all vulnerabilities in the source “Linux” package. 
		For example CVE-2019-9506
	\item\textbf{False Negatives by Vuls-} Vuls is missing some of the vulnerabilities that it should detect. 
		For example CVE-2019-5094
\end{enumerate}

\subsection{V-A}
V-A represents the vulnerabilities that are detected by Vuls but not detected by Anchore. Below are the reasons:
\begin{enumerate}
\item\textbf{EPOCH bug-} There is a bug in the Anchore logic due to which some vulnerabilities are not detected 
	by Anchore. This bug gets triggered when package’s version have epoch (1:3.3.9-1ubuntu2.3)  in it. 
		For example CVE-2018-1125 in procps package
\item\textbf{Ignored (out of standard support) bug-} There is another bug in Anchore due to which it is not 
	able to detect some vulnerabilities. Due to this bug Anchore does not detect vulnerabilities which 
		have a status as “ignored (out of standard support)” in Ubuntu 14.04 LTS. But these vulnerabilities 
		are detected by Vuls in Ubuntu 14.04 LTS. For example CVE-2019-13565 in Ubuntu 14.04
\item\textbf{With status as “Not-affected”-}  It means these vulnerabilities are affecting Ubuntu but they are not 
	affecting particular Ubuntu release that is present inside the image. The reason why its not affecting that 
		release is also mentioned. For example, one reason can be that code which is creating issue is not 
		present in that release etc. For example CVE-2015-2305 is not affecting package cups in Trusty.
\item\textbf{Rejected CVEs-} There are some CVE that are rejected because they are duplicate copies of another CVEs. 
	So Anchore is not reporting those rejected CVEs but Vuls is reporting those. For example CVE-2017-13753
	\end{enumerate}

\subsection{Effect Of Update}

The next experiment involves updating all these images and then rescanning them to see how the number of
vulnerabilities drop inside them. We used a \textit{bash} script to update these images. It will find
package manager inside the image and then update the image according the package manager it found.
When these images are rescanned after updation, results changed drastically. However, there are still
vulnerabilities reported by these scanners but there are no fixes available for these vulnerabilities
till date so there is nothing that can be done to remove these vulnerabilties. There are only two
options either you accept those vulnerabilities and use that image, or you discard that image.
So one thing is clear here that even if we keep our images up to date there will be still some
vulnerabilities that we have to face.
Below given table shows number of vulnerabilities that are detected by Anchore after updating these images.

\begin{figure}[!htb]
        {\includegraphics[width=\columnwidth]
        {Figures/afterupdate.png}}
        \caption{\label{fig:graph2} Scatterplot representing difference in number of vulnerabilities before and 
	after update of image}
      \end{figure}
\subsection{Effect Of Size Reduction}

\bibliography{bibliography}
